Propositions as Sessions, Mechanically

### 2015-04-03 paper/formalisation.tex
Added discussion of GV.

typing judgements

equivalences and axiomatic cut

principal cut rules

commuting conversions
\end{comment}

@@

The principal cut elimination rules are presented in Figure~\ref{??}. These
rules are expressed to indicate some additional complexity not present in the
original presentation. For example, structural equivalences and associativity
are not unique and different process terms result depending on initial binder
permutation. ${0 <~> 1}P$ indicates the permuting of binders $0$ and $1$ in
process P. The paper presentation specifies only one possible associativity
rule lending to the specific split of the linear environments. The reduction
rules were complicated to prove due to being composed of multiple constructor
instances and differed significantly to the standard operational semantics one
typically deals with in lambda calculus formalisms.

@@

@@

terms

typing judgements
\end{comment}

@@

While the aim is to be faithful to \citeauthor{Wadler:2014}'s description, I
may need to alter the approach slightly to avoid tricky formulations. For
instance, instead of providing $n$-ary branch and choice in GV, I provide
only binary versions of these operators to simplify the development. This
change is not restrictive however, since CP itself provides only binary
versions of plus ($\oplus$) and with ($\with$) constructs. Likewise, on a more
technical note, I amalgamate all types into one inductive definition and
provide a predicate for restricting typing rules to consider only valid
session types. \citeauthor{Wadler:2014} defines GV types as a set of mutually
recursive definitions; session types are types which may contain types as
subcomponents (arguments to send, for example). Unfortunately, handling
mutually inductive definitions in Coq can be quite involved; requiring one to
either rely on the Coq system to provide a stronger mutual induction principle
or defining one manually. Such a definition is possible, but it complicates
elimination via the induction principle.

another issue is the way i tried to formulate types initially treating them as
indexed families indexed by kinds. while this exploits the dependently typed
approach it complicated later definitions and proofs. for instance decidablity
of equality on types now relied on an axiom permitting heterogeneous
equality. further when formulating the typing environment we must decide which
kind each type had and most rules are polymorphic in that decision leading to
the creation of an awkward inspection function. these issues highlighted that
one should be careful and considerate how the representation will affect later
expressiveness.

>>>>

equivalences and axiomatic cut
\end{comment}

@@

The principal cut elimination rules are presented in Figure~\ref{??}. These
rules are expressed to indicate some additional complexity not present in the
original presentation. For example, structural equivalences and associativity
are not unique and different process terms result depending on initial binder
permutation. ${0 <~> 1}P$ indicates the permuting of binders $0$ and $1$ in
process P. The paper presentation specifies only one possible associativity
rule lending to the specific split of the linear environments. The reduction
rules were complicated to prove due to being composed of multiple constructor
instances and differed significantly to the standard operational semantics one
typically deals with in lambda calculus formalisms.
linear environments caused a problem with the splitting up of variables.

@@

The commuting conversions are shown in Figures~\ref{??} and~\ref{??}. Notice
that permutating binders is essential to formulating the majority of these
rules.

@@

terms
\end{comment}

@@

The typing judgements for the functional language GV are shown in
Figure~\ref{??}. Note that I have altered the $n$-ary branch and choice
constructs to the restricted binary versions in order to avoid tricky
formulations. The change is not restrictive however, since CP itself provides
only binary versions of plus ($\oplus$) and with ($\with$) constructs.

On a more technical note, all types are amalgamated into one inductive
definition and provide a predicate for restricting typing rules to consider
only valid session types. \citeauthor{Wadler:2014} defines GV types as a set
of mutually recursive definitions; session types are types which may contain
types as subcomponents (arguments to send, for example). Unfortunately,
handling mutually inductive definitions in Coq can be quite involved;
requiring one to either rely on the Coq system to provide a stronger mutual
induction principle or defining one manually. Such a definition is possible,
but it complicates applications of the induction principle.

Another issue related to types is how to representing type kinds (linearity or
non-linearity). Initially the type definition was defined as an indexed
family, indexed by the kind. While this definition has the advantage of
preventing certain ill-formed types, such as a non-linear session type, it
complicated later definitions and proofs. For instance, the decidability of
equality on types now relied on an axiom permitting heterogeneous
equality~\cite{??}. Generally, it is better to avoid unnecessary postulates if
possible. Further, when formulating the typing judgements the kind must be
given to each type hypothesised and most rules (e.g. ...) are polymorphic in
the kind, motivating an awkward inspection function to be applied per
application of a constructor \todo{clean this description up}. A solution to
this was to define a well-formed type relation and do away with the type
dependency. these issues highlighted that one should be careful and
considerate regarding when to use dependent types. these issues highlighted
that one should be careful and considerate how the representation will affect
later expressiveness.
